---
title: "Comoto Cover"
author: "Malia Crane"
date: "8/24/2021"
output: html_document
---

```{r Libraries,include=FALSE,message=FALSE}
library(tidyverse)
library(fpp3)
#Import Data, classify column data types
data <- read.csv("C:/Users/cran5028/Downloads/ComotoCrane-main/ecom_data.csv", row.names=NULL, stringsAsFactors=TRUE)
data$CustomerID<-as.factor(data$CustomerID)
data$InvoiceDay<-as.Date(data$InvoiceDay)
data$InvoiceDay<-as.POSIXct(data$InvoiceDay,format="%Y-%M-%D")
data$ID <- as.character(seq.int(nrow(data)))

#Filter out identified issues with data, negative sales amounts and SKU outliers:
data%>%filter(Sales >0)->data
data[!(data$SKU=="C2" | data$SKU=="BANK CHARGES" | data$SKU=="CRUK" | data$SKU=="D" | data$SKU=="DOT" | data$SKU=="M" | data$SKU=="PADS" | data$SKU=="POST"),]->data
droplevels(data$SKU)
```
### Overview
This file holds a brief overview and narrative responses for the technical prompt.

Other files include:

> ComotoCraneDescriptives - A review of the dataset and its characteristics.

> ComotoCraneTemporal - An analysis first creating Non-refunded orders in the 90 day window, followed by the full time series analysis. 

### Description of Sales Trends
```{r TotalSalesPlot,echo=FALSE,message=FALSE}
data%>%
     filter(Sales>0)%>%
     select(InvoiceDay,Sales)%>%
     arrange(InvoiceDay,Sales)%>%
     mutate(month=month(InvoiceDay),year=year(InvoiceDay))%>%
     group_by(year,month)%>%
     summarise(SumSales=sum(Sales))%>%
     arrange(year,month,SumSales)->tsdata


 tsdata$tsunit<-paste(paste(tsdata$year,"-"),tsdata$month)
 tsdata[,c(ncol(tsdata),1:(ncol(tsdata)-1))]->tsdata
 tsdata$year<-NULL
 tsdata$month<-NULL
 tsdata%>%arrange(tsunit)->tsdata
 
 tsdata%>%
     mutate(Month=yearmonth(tsunit))%>%
     as_tsibble(index=Month)->tsdata
 tsdata$tsunit<-NULL
 
 autoplot(tsdata,SumSales)+labs(title="Sum of Sales by Month",y="Sum of Sales",x="1 Month Scaling")
```

> Sales have been on an increasing volume during the period, with a marked increase between July and November 2011, ending with a dramatic decline during December 2011. 

```{r TotalSalesPlot2,echo=FALSE,message=FALSE}
data%>%
  filter(Sales>0)%>%
  select(InvoiceDay,Channel,Sales)%>%
  arrange(Channel,InvoiceDay,Sales)%>%
  mutate(month=month(InvoiceDay),year=year(InvoiceDay))%>%
  group_by(year,month,Channel)%>%
  summarise(SumSales=sum(Sales))%>%
  arrange(year,month,SumSales)->tsdata

#Making a better looking date column for visuals.
tsdata$tsunit<-paste(paste(tsdata$year,"-"),tsdata$month)
tsdata[,c(ncol(tsdata),1:(ncol(tsdata)-1))]->tsdata
tsdata$year<-NULL
tsdata$month<-NULL
tsdata%>%arrange(Channel,tsunit)->tsdata

tsdata%>%
mutate(Month=yearmonth(tsunit))%>%
as_tsibble(index=Month,key=Channel)->tsdata

#Remove extraneous column
tsdata$tsunit<-NULL

autoplot(tsdata,SumSales)+labs(title="Sum of Sales by Month and Channel",y="Sum of Sales",x="1 Month Scaling")
```

> This decline was experienced across all channels with the exception of Organic Social which saw an increase in sales. Forecasted sales estimates for the next 90 days show a gradual increase in sales to the prior year's observations, strongest for the Store channel.

> Customer metrics provide additional insights, with a typical Sales Order having a median of \$34.8 and a typical Customer purchase at \$32.5. During the period at a total sales level, most customers have a median sales purchase total of \$2.2k. 


### Recommendations to Increase Sales

> 1: Store Channel is associated with the highest total sum of sale profits, but is also associated with the greatest crash seen in nearly all channels during Nov/Dec 2011. The most stable path forward is in understanding the cause of this drop and formenting a contingency plan. Although not a top contributor, Organic Social has steady growth and is the sole increasing channel during the crash period, suggesting development in this area may bolster resiliency against these types of crashes.

> 2: Optimising suggested items/item prices to customers nearing \$35 (the median per-sale per customer) to generate slightly higher total per-sale costs (e.g. ~\$38). In a similar vein, customers were found to have an average of 15 and 22 items purchased per year/order (Median item quantity per period/order). A lower average quantity per period than quantity per order indicates that undocumented issues may exist with quality/delivery of services and that customers may be not be returning. 
